---
title: "NDBC-rnoaa"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MUR Sea Surface Temperature at points

This tool will harvest [NDBC](https://www.ndbc.noaa.gov/) buoy using the [rnoaa](https://docs.ropensci.org/rnoaa/) R package from [rOpenSci](https://ropensci.org/).

#### Setup

Have the file `NDBC-rnoaa.R` in your working directory

#### Required packages

If you don't have the [rnoaa](https://CRAN.R-project.org/package=rnoaa), [dplyr](https://CRAN.R-project.org/package=dplyr) and [readr](https://CRAN.R-project.org/package=readr) packages installed then they will be installed for you automatically (from CRAN).

#### Source

Source the file: (be sure it is in your working directory)

```{r source, message = FALSE}
source('NDBC-rnoaa.R')
```

#### Know your buoy

See the [tutorial](../tutorials/NDBC-buoydata) for determining your buoy's `id` and the datasets available to it.

### Get the buoy data

```{r}
x <- get_buoy_data(dataset = 'stdmet',
                   buoyid = "CFWM1",
                   years = c(2014, 2017:2018),
                   outfile = "~/CFWM1-stdmet.csv.gz")
```
And plot...

```{r, warning = FALSE, message = FALSE}
ggplot2::ggplot(data = x, ggplot2::aes(x = time, y = sea_surface_temperature)) +
  ggplot2::geom_point(size = 0.2, alpha = 0.4) + 
  ggplot2::labs(title = "CFWM1-stdmet") 
```

---

Developer notes:

 - This is quick-and-dirty
 
 - There is limited error checking, so use at your own risk
 